Adaptive Approximation for Multivariate Linear Problems with Inputs Lying in a Cone

### 2019-10-31 DingHickKritzMakRICAMChapVerF.tex
Updates from Overleaf

\DeclareMathOperator{\DHKMSOL}{SOL}
\DeclareMathOperator{\DHKMAPP}{APP}
\DeclareMathOperator{\DHKMALG}{ALG}
\DeclareMathOperator{\DHKMERR}{ERR}
\DeclareMathOperator{\DHKMcard}{card}
\newcommand{\DHKMdataN}{\bigl(\hf(\bsk_i)\bigr)_{i=1}^n}
\newcommand{\DHKMdataNj}{\bigl(\hf(\bsk_i)\bigr)_{i=1}^{n_j}}
\newcommand{\DHKMdataNjd}{\bigl(\hf(\bsk_i)\bigr)_{i=1}^{n_{j^\dagger}}}
\newcommand{\DHKMERRN}{\DHKMERR\bigl(\DHKMdataN,n\bigr)}
\newcommand{\DHKMERRNj}{\DHKMERR\bigl(\DHKMdataNj,n_j\bigr)}
\newcommand{\DHKMERRNjd}{\DHKMERR\bigl(\DHKMdataNjd,n_{j^\dagger}\bigr)}
\DeclareMathOperator{\DHKMCOST}{COST}
\DeclareMathOperator{\COMP}{COMP}
\newcommand{\hf}{\widehat{f}}
\newcommand{\hg}{\widehat{g}}
\newcommand{\tu}{\Tilde{u}}
\newcommand{\tv}{\Tilde{v}}
\newcommand{\tcalJ}{\widetilde{\calJ}}
\newcommand{\tcalK}{\widetilde{\calK}}
\newcommand{\tlambda}{\Tilde{\lambda}}
\newcommand{\tbbK}{\widetilde{\bbK}}
\newcommand{\lo}{\textup{lo}}
\newcommand{\up}{\textup{up}}
\newcommand{\inc}{\textup{in}}
\newcommand{\out}{\textup{out}}
\newcommand{\E}{\textup{e}}
\def\abs#1{\ensuremath{\left \lvert #1 \right \rvert}}
\newcommand{\normabs}[1]{\ensuremath{\lvert #1 \rvert}}
\newcommand{\bigabs}[1]{\ensuremath{\bigl \lvert #1 \bigr \rvert}}
\newcommand{\Bigabs}[1]{\ensuremath{\Bigl \lvert #1 \Bigr \rvert}}
\newcommand{\biggabs}[1]{\ensuremath{\biggl \lvert #1 \biggr \rvert}}
\newcommand{\Biggabs}[1]{\ensuremath{\Biggl \lvert #1 \Biggr \rvert}}
\newcommand{\norm}[2][{}]{\ensuremath{\left \lVert #2 \right \rVert}_{#1}}
\newcommand{\normnorm}[2][{}]{\ensuremath{\lVert #2 \rVert}_{#1}}
\newcommand{\bignorm}[2][{}]{\ensuremath{\bigl \lVert #2 \bigr \rVert}_{#1}}
\newcommand{\Bignorm}[2][{}]{\ensuremath{\Bigl \lVert #2 \Bigr \rVert}_{#1}}
\newcommand{\biggnorm}[2][{}]{\ensuremath{\biggl \lVert #2 \biggr \rVert}_{#1}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}
\newcommand{\FredNote}[1]{{\color{blue}Fred: #1}}
\newcommand{\YuhanNote}[1]{{\color{magenta}Yuhan: #1}}
\newcommand{\PeterNote}[1]{{\color{orange}Peter: #1}}
\newcommand{\SimonNote}[1]{{\color{purple}Simon: #1}}
\newcommand{\DHKMchange}[1]{{\color{red}{#1}}}

@@

We assume in this chapter that the functions $f\in\calF$ can be represented in terms of a basis representation as $f = \sum_{\bsk \in \mathbb{K}} \hf(\bsk) u_{\bsk}$, and we allow algorithms $\DHKMALG (f,\varepsilon)$ to use $n(f,\varepsilon)$ coefficients $\hf (\bsk)$ of this representation. That is, 
we study algorithms that have an output of the form 
\begin{equation*} %\label{DHKM:algform}
  \DHKMALG(f,\varepsilon) = \sum_{i=1}^{n(f,\varepsilon)}\hf(\bsk_i) v_{\bsk_i} \qquad \forall f \in \calF, \ n(f,\varepsilon) \in \N,
\end{equation*}
where an oracle provides $\hf(\bsk_1),\ldots,\hf (\bsk_{n(f,\varepsilon)})$ 
with $n(f,\varepsilon)$ being chosen based on a stopping rule. 
We assume that it is possible to sample the series coefficients of the input function.  This is a more restrictive assumption than being able to sample any linear functional, but it is different than only being able to sample function values.  An important future problem is to extend the theory in this chapter to the case where the only function data available are function values.

@@

A crucial concept is the set of all possible algorithms that may be constructed using series coefficients and that \emph{satisfy error criterion \eqref{DHKM:err_crit}}. We denote this set by $\calA(\calC)$. 
Furthermore, we define the \emph{computational complexity} of a problem as the information cost of the best algorithm:
\begin{equation*}
    \COMP(\calC, \varepsilon,R) := \min_{\DHKMALG \in \calA(\calC)} \DHKMCOST(\DHKMALG, \calC, \varepsilon,R) .
\end{equation*}
These definitions follow the information-based complexity literature \citep{TraWer98, TraWasWoz88}.
We define an algorithm to be \emph{essentially optimal} if there exist some fixed positive $\omega$, $\varepsilon_{\max}$, and $R_{\min}$ for which
\begin{multline} \label{DHKM:EssentialOpt}
    \DHKMCOST(\DHKMALG, \calC, \varepsilon,R) \le \COMP(\calC, \omega \varepsilon,R) \\ \forall \, \varepsilon \in (0, \varepsilon_{\max}], \ R \in [R_{\min}, \infty).
\end{multline}
If the complexity of the problem is $\calO(\varepsilon^{-p})$, the cost of an essentially optimal algorithm is also $\calO(\varepsilon^{-p})$. If the complexity of the problem is asymptotically $c \log(1 + \varepsilon^{-1})$, then  the cost of an essentially optimal algorithm is also asymptotically $c \log(1 + \varepsilon^{-1})$. 
We will show that our adaptive algorithms presented in Sections \ref{DHKM:pilot_sec} and \ref{DHKM:tracking_sec}  are essentially optimal.

@@

The first definition of a cone is given in \eqref{DHKM:pilot_cone}, 
where the idea is that the cone consists of $f\in\calF$ such that 
$$
  \norm[\calF]{f}\le h (\hf(\bsk_1),\ldots,\hf (\bsk_{n_1}))
$$
for a suitably chosen homogeneous function $h:\bbR^{n_1}\to [0,\infty)$. This means that the norm of $f$ can be bounded well in terms of 
a \emph{pilot sample} of size $n_1\in\bbN$. See \eqref{DHKM:pilot_cone} for the precise definition. In Section \ref{DHKM:SecAdapAlgTrackDecay}, we will modify the definition of cones such that we can keep track of the decay rate of the series coefficients of the input function (see \eqref{DHKM:TrackConeDef}), 
and in Section \ref{DHKM:smoothimportance_sec} we allow a situation 
where we infer coordinate and smoothness importance from some initial data, thus making the definition of the cone more practical (see  
\eqref{DHKM:pilot_cone2}).

@@

Now, we define our problem more formally.  The solution operator $\DHKMSOL$ maps the input function to an output, $\DHKMSOL:\calF \to \calG$.  The Banach spaces of inputs and outputs are defined by series expansions:
\begin{gather*}
    \calF := \left \{f = \sum_{\bsk \in \mathbb{K}} \hf(\bsk) u_{\bsk} : \norm[\calF]{f} : = \norm[\rho]{\left( \frac{\hf(\bsk)}{\lambda_{\bsk}} \right)_{\bsk \in \mathbb{K}}} < \infty \right\}, \quad\mbox{with}\quad 1 \le \rho \le \infty, 
\end{gather*}
and
\begin{gather*}
    \calG := \left \{g = \sum_{\bsk \in \mathbb{K}} \hg(\bsk) v_{\bsk} : \norm[\calG]{g} : = \norm[\tau]{\bigl(  \hg(\bsk)  \bigr)_{\bsk \in \mathbb{K}}} < \infty \right\}, \quad\mbox{with}\quad 1 \le \tau \le \rho.
\end{gather*}

@@

The $\lambda_{\bsk}$ represent the importance of the series coefficients of the input function.  The larger $\lambda_{\bsk}$ is, the more important $\hf(\bsk)$ is.  The $\lambda_{\bsk}$ are assumed to have a known order:
\begin{equation} \label{DHKM:lambda_order}
    \lambda_{\bsk_1} \ge \lambda_{\bsk_2} \ge \cdots >0,
\end{equation}
implies an ordering of the wavenumbers, $\bsk$. We also assume that all $\lambda_{\bsk}$ are  positive to avoid the trivial case where $\DHKMSOL(f)$ can be expressed exactly as a finite sum for all $f \in \calF$.  We denote the sequence of weights  $\bigl(\lambda_{\bsk_i}\bigr)_{i=1}^\infty$ by $\bslambda$.

@@

\begin{lemma} \label{DHKM:Key_Lem}
Let $\calK$ be some proper or improper subset of the index set $\bbK$. Moreover, let $\rho'$ be defined by the relation
\begin{equation*}
    \frac 1\rho + \frac 1 {\rho'} = \frac 1 \tau, \qquad \text{i.e., } \rho' := \frac{\rho \tau}{\rho - \tau},
\end{equation*}
so $\tau \le \rho' \le \infty$.  Let $\Lambda :=  \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \calK}}$ be the norm of a subset of the weights.  Then the following are true for $f = \sum_{\bsk \in \calK} \hf(\bsk) u_{\bsk}$:
\begin{equation}
\label{DHKM:SOL_ineq}
    \norm[\calG]{\DHKMSOL(f)} = \norm[\tau]{\bigl(\hf(\bsk) \bigr)_{\bsk \in \calK}} \le \norm[\calF]{f} \, \Lambda,
    \end{equation}
    \begin{multline}
    \label{DHKM:SOL_tight_ineq}
    \bigabs{\hf(\bsk)} = \begin{cases}
    \displaystyle 
    \frac{R \lambda_{\bsk}^{\rho'/\rho + 1}}{\Lambda^{\rho'/\rho}}, & \forall \bsk \in \calK, \quad  \text{if}\ \rho'<\infty, \\
    R \Lambda \delta_{\bsk,\bsk^*}, & \forall \bsk \in \calK, \ \bsk^* \in \calK \text{ satisfies } \lambda_{\bsk^*} = \Lambda, \quad\text{if}\ \rho' = \infty,
    \end{cases}
   \\ 
    \implies  \ \ \norm[\calF]{f} = R \ \mbox{and} \ \norm[\calG]{\DHKMSOL(f)} = R \Lambda.
    \end{multline}
Equality \eqref{DHKM:SOL_tight_ineq} illustrates how inequality \eqref{DHKM:SOL_ineq} may be made tight.
\end{lemma}
\begin{proof}
We give the proof for $\rho' < \infty$.  The proof for $\rho' = \infty$ follows similarly. 
The proof of inequality \eqref{DHKM:SOL_ineq} proceeds by applying H\"older's inequality:  
\begin{align}
    \label{DHKM:SOL_A}
    \norm[\calG]{\DHKMSOL(f)}  
    & = \biggnorm[\calG]{\sum_{\bsk \in \calK} \hf(\bsk) v_{\bsk}} = \norm[\tau]{\bigl(  \hf(\bsk)  \bigr)_{\bsk \in \calK}}  = \left [\sum_{\bsk \in \calK}  \left\lvert\frac{\hf(\bsk)}{\lambda_{\bsk}} \right\rvert^{\tau} \lambda_{\bsk}^{\tau} \right]^{1/\tau} \\
    \nonumber
    & \le \biggnorm[\rho]{\biggl(  \frac{\hf(\bsk)}{\lambda_{\bsk}}  \biggr)_{\bsk \in \calK}} \, \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \calK}} = \norm[\calF]{f} \, \Lambda \qquad \text{since }\frac 1\rho + \frac 1 {\rho'} = \frac 1 \tau.
\end{align}
Substituting the formula for $\bigabs{\hf(\bsk)}$ in \eqref{DHKM:SOL_tight_ineq} into equation \eqref{DHKM:SOL_A} and applying the relationship between $\rho$, $\rho'$, and $\tau$ yields
\begin{equation*}
       \norm[\calG]{\DHKMSOL(f)}  
    =  \frac{R \bignorm[\tau]{\bigl(  \lambda_{\bsk}^{\rho'/\rho + 1}  \bigr)_{\bsk \in \calK}}} {\Lambda^{\rho'/\rho}} 
    = \frac{R \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \calK}}^{\rho'/\rho + 1}}
    {\Lambda^{\rho'/\rho}} = R \Lambda.
\end{equation*}
Moreover,
\begin{equation*}
    \norm[\calF]{f}  
    = \norm[\rho]{\left( \frac{\hf(\bsk)}{\lambda_{\bsk}} \right)_{\bsk \in \calK}}
    = \frac{R \bignorm[\rho]{\bigl(  \lambda_{\bsk}^{\rho'/\rho}  \bigr)_{\bsk \in \calK}}}{\Lambda^{\rho'/\rho}} 
    = \frac{R \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \calK}}^{\rho'/\rho}}
    {\Lambda^{\rho'/\rho}} = R.
\end{equation*}
This completes the proof.
\end{proof} \

@@

Consider the case of approximating functions defined over $[-1,1]^d$, using a Chebyshev polynomial basis.  The input function is denoted $f$, and the solution is $\DHKMSOL(f) = f$.  In this case,
	\begin{align*}
	f &  = \sum_{\bsk \in \bbN_0^d} \widehat{f}(\bsk) u_\bsk =: \DHKMSOL(f), 
	\qquad \bsk = (k_1, \ldots, k_d) \in \N_0^d,\\
    u_\bsk & := \prod_{\ell =1}^d \tu_{k_\ell} , 
    \qquad \tu_{k}(x) := \cos( k \cos^{-1}(x)) \quad \forall k \in \N_0.
	\end{align*}
For this example, let $\calF$ denote a Hilbert space of input functions where we choose $\bbK=\bbN_0^d$ and $\rho=2$,
\begin{equation}
\label{DHKM:IllExWts}
    \norm[\calF]{f} := \norm[2]{\left(\frac{\hf(\bsk)}{\lambda_{\bsk}}\right)_{\bsk \in \N_0^d}}, \qquad \text{where } \lambda_\bsk := \prod_{\substack{\ell =1\\ k_\ell > 0}}^d \frac{w_\ell}{k^r_\ell}, \quad r > 0.
\end{equation}
The $w_\ell$ are non-negative \emph{coordinate weights}, which embody the assumption that $f$ may depend more strongly on coordinates with larger $w_\ell$ than those with smaller $w_\ell$.  The definition of the $\calF$-norm implies that an input function must have series coefficients that decay quickly enough as the degree of the polynomial increases.  Larger $r$ implies smoother input functions.

Define the norm of the Hilbert space of outputs as
\begin{equation*}
      \norm[\calG]{\sum_{\bsk \in \N_0^d} \hg(\bsk) u_{\bsk}} : = \norm[2]{\bigl(\hg(\bsk) \bigr)_{\bsk \in \N_0^d}}.
\end{equation*}
Due to the ordering of the weights in \eqref{DHKM:lambda_order}, it is natural to approximate the solution using the first $n$ series coefficients as follows:
\begin{equation*} %\label{DHKM:Ex_APP_def}
   \DHKMAPP(f,n) := \sum_{i=1}^n \hf(\bsk_i) u_{\bsk_i} \qquad \forall f \in \calF, \ n \in \N.
\end{equation*}
The error of this approximation in terms of the norm on the output space, $\calG$, can be expressed as
\begin{equation*}
    \norm[\calG]{\DHKMSOL(f) - \DHKMAPP(f,n)} = \norm[2]{\bigl(\hf(\bsk_i) \bigr)_{i=n+1}^{\infty}}.
\end{equation*}
Here we have chosen $\tau = 2$ so $\rho'=\infty$.  If one has a fixed data budget, $n$, then $\DHKMAPP(f,n)$ is the best answer for certain classes of inputs in $\calF$ as for example balls, which are among the most prominent examples studied in the literature.

@@

    \RETURN $\DHKMALG(f,\varepsilon) = \DHKMAPP(f,n^*)$
\end{algorithmic}
\end{algorithm}
We now show that this algorithm is optimal if the input set is a ball.
\begin{theorem}\label{DHKM:thm_cost_non_adapt}
The non-adaptive Algorithm \ref{DHKM:BallAlg} has an information cost for the set of input functions $\calC = \calB_R$ that is given by
\[
\DHKMCOST(\DHKMALG, \calB_R, \varepsilon,R') = \min \left \{n \in \N_0 : \bignorm[\rho']{\bslambda^{(n+1)}} \le \varepsilon /R \right \}, \qquad R’ \in (0, R].
\]
This algorithm is essentially optimal for the set of input functions $\calB_R$, namely,
\begin{multline*}
\DHKMCOST(\DHKMALG, \calB_R, \varepsilon,R') \le \COMP(\calB_R,\omega \varepsilon ,R') \\ \forall \, \varepsilon \in (0, \varepsilon_{\max}], \ R’ \in  [ R_{\min} , R], \ R \in [R_{\min}, \infty), 
\end{multline*}
where $\varepsilon_{\max}$ and $R_{\min}$ are arbitrary and fixed, and $\omega = R_{\min}/R$.
\end{theorem}

>>>>

\DeclareMathOperator{\DHKMSOL}{SOL}
\DeclareMathOperator{\DHKMAPP}{APP}
\DeclareMathOperator{\DHKMALG}{ALG}
\DeclareMathOperator{\DHKMERR}{ERR}
\DeclareMathOperator{\DHKMcard}{card}
\newcommand{\DHKMdataN}{\bigl(\DHKMhf(\bsk_i)\bigr)_{i=1}^n}
\newcommand{\DHKMdataNj}{\bigl(\DHKMhf(\bsk_i)\bigr)_{i=1}^{n_j}}
\newcommand{\DHKMdataNjd}{\bigl(\DHKMhf(\bsk_i)\bigr)_{i=1}^{n_{j^\dagger}}}
\newcommand{\DHKMERRN}{\DHKMERR\bigl(\DHKMdataN,n\bigr)}
\newcommand{\DHKMERRNj}{\DHKMERR\bigl(\DHKMdataNj,n_j\bigr)}
\newcommand{\DHKMERRNjd}{\DHKMERR\bigl(\DHKMdataNjd,n_{j^\dagger}\bigr)}
\DeclareMathOperator{\DHKMCOST}{COST}
\DeclareMathOperator{\DHKMCOMP}{COMP}
\newcommand{\DHKMhf}{\widehat{f}}
\newcommand{\hg}{\widehat{g}}
\newcommand{\tu}{\Tilde{u}}
\newcommand{\tv}{\Tilde{v}}
\newcommand{\tcalJ}{\widetilde{\calJ}}
\newcommand{\tcalK}{\widetilde{\calK}}
\newcommand{\tlambda}{\Tilde{\lambda}}
\newcommand{\tbbK}{\widetilde{\bbK}}
\newcommand{\lo}{\textup{lo}}
\newcommand{\up}{\textup{up}}
\newcommand{\inc}{\textup{in}}
\newcommand{\out}{\textup{out}}
\newcommand{\E}{\textup{e}}
\def\abs#1{\ensuremath{\left \lvert #1 \right \rvert}}
\newcommand{\normabs}[1]{\ensuremath{\lvert #1 \rvert}}
\newcommand{\bigabs}[1]{\ensuremath{\bigl \lvert #1 \bigr \rvert}}
\newcommand{\Bigabs}[1]{\ensuremath{\Bigl \lvert #1 \Bigr \rvert}}
\newcommand{\biggabs}[1]{\ensuremath{\biggl \lvert #1 \biggr \rvert}}
\newcommand{\Biggabs}[1]{\ensuremath{\Biggl \lvert #1 \Biggr \rvert}}
\newcommand{\norm}[2][{}]{\ensuremath{\left \lVert #2 \right \rVert}_{#1}}
\newcommand{\normnorm}[2][{}]{\ensuremath{\lVert #2 \rVert}_{#1}}
\newcommand{\bignorm}[2][{}]{\ensuremath{\bigl \lVert #2 \bigr \rVert}_{#1}}
\newcommand{\Bignorm}[2][{}]{\ensuremath{\Bigl \lVert #2 \Bigr \rVert}_{#1}}
\newcommand{\biggnorm}[2][{}]{\ensuremath{\biggl \lVert #2 \biggr \rVert}_{#1}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}
\newcommand{\FredNote}[1]{{\color{blue}Fred: #1}}
\newcommand{\YuhanNote}[1]{{\color{magenta}Yuhan: #1}}
\newcommand{\PeterNote}[1]{{\color{orange}Peter: #1}}
\newcommand{\SimonNote}[1]{{\color{purple}Simon: #1}}
\newcommand{\DHKMchange}[1]{{\color{red}{#1}}}

@@

We assume in this chapter that the functions $f\in\calF$ can be represented in terms of a basis representation as $f = \sum_{\bsk \in \mathbb{K}} \DHKMhf(\bsk) u_{\bsk}$, and we allow algorithms $\DHKMALG (f,\varepsilon)$ to use $n(f,\varepsilon)$ coefficients $\DHKMhf (\bsk)$ of this representation. That is, 
we study algorithms that have an output of the form 
\begin{equation*} %\label{DHKM:algform}
  \DHKMALG(f,\varepsilon) = \sum_{i=1}^{n(f,\varepsilon)}\DHKMhf(\bsk_i) v_{\bsk_i} \qquad \forall f \in \calF, \ n(f,\varepsilon) \in \N,
\end{equation*}
where an oracle provides $\DHKMhf(\bsk_1),\ldots,\DHKMhf (\bsk_{n(f,\varepsilon)})$ 
with $n(f,\varepsilon)$ being chosen based on a stopping rule. 
We assume that it is possible to sample the series coefficients of the input function.  This is a more restrictive assumption than being able to sample any linear functional, but it is different than only being able to sample function values.  An important future problem is to extend the theory in this chapter to the case where the only function data available are function values.

@@

A crucial concept is the set of all possible algorithms that may be constructed using series coefficients and that \emph{satisfy error criterion \eqref{DHKM:err_crit}}. We denote this set by $\calA(\calC)$. 
Furthermore, we define the \emph{computational complexity} of a problem as the information cost of the best algorithm:
\begin{equation*}
    \DHKMCOMP(\calC, \varepsilon,R) := \min_{\DHKMALG \in \calA(\calC)} \DHKMCOST(\DHKMALG, \calC, \varepsilon,R) .
\end{equation*}
These definitions follow the information-based complexity literature \citep{TraWer98, TraWasWoz88}.
We define an algorithm to be \emph{essentially optimal} if there exist some fixed positive $\omega$, $\varepsilon_{\max}$, and $R_{\min}$ for which
\begin{multline} \label{DHKM:EssentialOpt}
    \DHKMCOST(\DHKMALG, \calC, \varepsilon,R) \le \DHKMCOMP(\calC, \omega \varepsilon,R) \\ \forall \, \varepsilon \in (0, \varepsilon_{\max}], \ R \in [R_{\min}, \infty).
\end{multline}
If the complexity of the problem is $\calO(\varepsilon^{-p})$, the cost of an essentially optimal algorithm is also $\calO(\varepsilon^{-p})$. If the complexity of the problem is asymptotically $c \log(1 + \varepsilon^{-1})$, then  the cost of an essentially optimal algorithm is also asymptotically $c \log(1 + \varepsilon^{-1})$. 
We will show that our adaptive algorithms presented in Sections \ref{DHKM:pilot_sec} and \ref{DHKM:tracking_sec}  are essentially optimal.

@@

The first definition of a cone is given in \eqref{DHKM:pilot_cone}, 
where the idea is that the cone consists of $f\in\calF$ such that 
$$
  \norm[\calF]{f}\le h (\DHKMhf(\bsk_1),\ldots,\DHKMhf (\bsk_{n_1}))
$$
for a suitably chosen homogeneous function $h:\bbR^{n_1}\to [0,\infty)$. This means that the norm of $f$ can be bounded well in terms of 
a \emph{pilot sample} of size $n_1\in\bbN$. See \eqref{DHKM:pilot_cone} for the precise definition. In Section \ref{DHKM:SecAdapAlgTrackDecay}, we will modify the definition of cones such that we can keep track of the decay rate of the series coefficients of the input function (see \eqref{DHKM:TrackConeDef}), 
and in Section \ref{DHKM:smoothimportance_sec} we allow a situation 
where we infer coordinate and smoothness importance from some initial data, thus making the definition of the cone more practical (see  
\eqref{DHKM:pilot_cone2}).

@@

Now, we define our problem more formally.  The solution operator $\DHKMSOL$ maps the input function to an output, $\DHKMSOL:\calF \to \calG$.  The Banach spaces of inputs and outputs are defined by series expansions:
\begin{gather*}
    \calF := \left \{f = \sum_{\bsk \in \mathbb{K}} \DHKMhf(\bsk) u_{\bsk} : \norm[\calF]{f} : = \norm[\rho]{\left( \frac{\DHKMhf(\bsk)}{\lambda_{\bsk}} \right)_{\bsk \in \mathbb{K}}} < \infty \right\}, \quad\mbox{with}\quad 1 \le \rho \le \infty, 
\end{gather*}
and
\begin{gather*}
    \calG := \left \{g = \sum_{\bsk \in \mathbb{K}} \hg(\bsk) v_{\bsk} : \norm[\calG]{g} : = \norm[\tau]{\bigl(  \hg(\bsk)  \bigr)_{\bsk \in \mathbb{K}}} < \infty \right\}, \quad\mbox{with}\quad 1 \le \tau \le \rho.
\end{gather*}

@@

The $\lambda_{\bsk}$ represent the importance of the series coefficients of the input function.  The larger $\lambda_{\bsk}$ is, the more important $\DHKMhf(\bsk)$ is.  The $\lambda_{\bsk}$ are assumed to have a known order:
\begin{equation} \label{DHKM:lambda_order}
    \lambda_{\bsk_1} \ge \lambda_{\bsk_2} \ge \cdots >0,
\end{equation}
implies an ordering of the wavenumbers, $\bsk$. We also assume that all $\lambda_{\bsk}$ are  positive to avoid the trivial case where $\DHKMSOL(f)$ can be expressed exactly as a finite sum for all $f \in \calF$.  We denote the sequence of weights  $\bigl(\lambda_{\bsk_i}\bigr)_{i=1}^\infty$ by $\bslambda$.

@@

\begin{lemma} \label{DHKM:Key_Lem}
Let $\calK$ be some proper or improper subset of the index set $\bbK$. Moreover, let $\rho'$ be defined by the relation
\begin{equation*}
    \frac 1\rho + \frac 1 {\rho'} = \frac 1 \tau, \qquad \text{i.e., } \rho' := \frac{\rho \tau}{\rho - \tau},
\end{equation*}
so $\tau \le \rho' \le \infty$.  Let $\Lambda :=  \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \calK}}$ be the norm of a subset of the weights.  Then the following are true for $f = \sum_{\bsk \in \calK} \DHKMhf(\bsk) u_{\bsk}$:
\begin{equation}
\label{DHKM:SOL_ineq}
    \norm[\calG]{\DHKMSOL(f)} = \norm[\tau]{\bigl(\DHKMhf(\bsk) \bigr)_{\bsk \in \calK}} \le \norm[\calF]{f} \, \Lambda,
    \end{equation}
    \begin{multline}
    \label{DHKM:SOL_tight_ineq}
    \bigabs{\DHKMhf(\bsk)} = \begin{cases}
    \displaystyle 
    \frac{R \lambda_{\bsk}^{\rho'/\rho + 1}}{\Lambda^{\rho'/\rho}}, & \forall \bsk \in \calK, \quad  \text{if}\ \rho'<\infty, \\
    R \Lambda \delta_{\bsk,\bsk^*}, & \forall \bsk \in \calK, \ \bsk^* \in \calK \text{ satisfies } \lambda_{\bsk^*} = \Lambda, \quad\text{if}\ \rho' = \infty,
    \end{cases}
   \\ 
    \implies  \ \ \norm[\calF]{f} = R \ \mbox{and} \ \norm[\calG]{\DHKMSOL(f)} = R \Lambda.
    \end{multline}
Equality \eqref{DHKM:SOL_tight_ineq} illustrates how inequality \eqref{DHKM:SOL_ineq} may be made tight.
\end{lemma}
\begin{proof}
We give the proof for $\rho' < \infty$.  The proof for $\rho' = \infty$ follows similarly. 
The proof of inequality \eqref{DHKM:SOL_ineq} proceeds by applying H\"older's inequality:  
\begin{align}
    \label{DHKM:SOL_A}
    \norm[\calG]{\DHKMSOL(f)}  
    & = \biggnorm[\calG]{\sum_{\bsk \in \calK} \DHKMhf(\bsk) v_{\bsk}} = \norm[\tau]{\bigl(  \DHKMhf(\bsk)  \bigr)_{\bsk \in \calK}}  = \left [\sum_{\bsk \in \calK}  \left\lvert\frac{\DHKMhf(\bsk)}{\lambda_{\bsk}} \right\rvert^{\tau} \lambda_{\bsk}^{\tau} \right]^{1/\tau} \\
    \nonumber
    & \le \biggnorm[\rho]{\biggl(  \frac{\DHKMhf(\bsk)}{\lambda_{\bsk}}  \biggr)_{\bsk \in \calK}} \, \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \calK}} = \norm[\calF]{f} \, \Lambda \qquad \text{since }\frac 1\rho + \frac 1 {\rho'} = \frac 1 \tau.
\end{align}
Substituting the formula for $\bigabs{\DHKMhf(\bsk)}$ in \eqref{DHKM:SOL_tight_ineq} into equation \eqref{DHKM:SOL_A} and applying the relationship between $\rho$, $\rho'$, and $\tau$ yields
\begin{equation*}
       \norm[\calG]{\DHKMSOL(f)}  
    =  \frac{R \bignorm[\tau]{\bigl(  \lambda_{\bsk}^{\rho'/\rho + 1}  \bigr)_{\bsk \in \calK}}} {\Lambda^{\rho'/\rho}} 
    = \frac{R \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \calK}}^{\rho'/\rho + 1}}
    {\Lambda^{\rho'/\rho}} = R \Lambda.
\end{equation*}
Moreover,
\begin{equation*}
    \norm[\calF]{f}  
    = \norm[\rho]{\left( \frac{\DHKMhf(\bsk)}{\lambda_{\bsk}} \right)_{\bsk \in \calK}}
    = \frac{R \bignorm[\rho]{\bigl(  \lambda_{\bsk}^{\rho'/\rho}  \bigr)_{\bsk \in \calK}}}{\Lambda^{\rho'/\rho}} 
    = \frac{R \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \calK}}^{\rho'/\rho}}
    {\Lambda^{\rho'/\rho}} = R.
\end{equation*}
This completes the proof.
\end{proof} \

@@

Consider the case of approximating functions defined over $[-1,1]^d$, using a Chebyshev polynomial basis.  The input function is denoted $f$, and the solution is $\DHKMSOL(f) = f$.  In this case,
	\begin{align*}
	f &  = \sum_{\bsk \in \bbN_0^d} \widehat{f}(\bsk) u_\bsk =: \DHKMSOL(f), 
	\qquad \bsk = (k_1, \ldots, k_d) \in \N_0^d,\\
    u_\bsk & := \prod_{\ell =1}^d \tu_{k_\ell} , 
    \qquad \tu_{k}(x) := \cos( k \cos^{-1}(x)) \quad \forall k \in \N_0.
	\end{align*}
For this example, let $\calF$ denote a Hilbert space of input functions where we choose $\bbK=\bbN_0^d$ and $\rho=2$,
\begin{equation}
\label{DHKM:IllExWts}
    \norm[\calF]{f} := \norm[2]{\left(\frac{\DHKMhf(\bsk)}{\lambda_{\bsk}}\right)_{\bsk \in \N_0^d}}, \qquad \text{where } \lambda_\bsk := \prod_{\substack{\ell =1\\ k_\ell > 0}}^d \frac{w_\ell}{k^r_\ell}, \quad r > 0.
\end{equation}
The $w_\ell$ are non-negative \emph{coordinate weights}, which embody the assumption that $f$ may depend more strongly on coordinates with larger $w_\ell$ than those with smaller $w_\ell$.  The definition of the $\calF$-norm implies that an input function must have series coefficients that decay quickly enough as the degree of the polynomial increases.  Larger $r$ implies smoother input functions.

Define the norm of the Hilbert space of outputs as
\begin{equation*}
      \norm[\calG]{\sum_{\bsk \in \N_0^d} \hg(\bsk) u_{\bsk}} : = \norm[2]{\bigl(\hg(\bsk) \bigr)_{\bsk \in \N_0^d}}.
\end{equation*}
Due to the ordering of the weights in \eqref{DHKM:lambda_order}, it is natural to approximate the solution using the first $n$ series coefficients as follows:
\begin{equation*} %\label{DHKM:Ex_APP_def}
   \DHKMAPP(f,n) := \sum_{i=1}^n \DHKMhf(\bsk_i) u_{\bsk_i} \qquad \forall f \in \calF, \ n \in \N.
\end{equation*}
The error of this approximation in terms of the norm on the output space, $\calG$, can be expressed as
\begin{equation*}
    \norm[\calG]{\DHKMSOL(f) - \DHKMAPP(f,n)} = \norm[2]{\bigl(\DHKMhf(\bsk_i) \bigr)_{i=n+1}^{\infty}}.
\end{equation*}
Here we have chosen $\tau = 2$ so $\rho'=\infty$.  If one has a fixed data budget, $n$, then $\DHKMAPP(f,n)$ is the best answer for certain classes of inputs in $\calF$ as for example balls, which are among the most prominent examples studied in the literature.

@@

    \RETURN $\DHKMALG(f,\varepsilon) = \DHKMAPP(f,n^*)$
\end{algorithmic}
\end{algorithm}
We now show that this algorithm is optimal if the input set is a ball.
\begin{theorem}\label{DHKM:thm_cost_non_adapt}
The non-adaptive Algorithm \ref{DHKM:BallAlg} has an information cost for the set of input functions $\calC = \calB_R$ that is given by
\[
\DHKMCOST(\DHKMALG, \calB_R, \varepsilon,R') = \min \left \{n \in \N_0 : \bignorm[\rho']{\bslambda^{(n+1)}} \le \varepsilon /R \right \}, \qquad R’ \in (0, R].
\]
This algorithm is essentially optimal for the set of input functions $\calB_R$, namely,
\begin{multline*}
\DHKMCOST(\DHKMALG, \calB_R, \varepsilon,R') \le \DHKMCOMP(\calB_R,\omega \varepsilon ,R') \\ \forall \, \varepsilon \in (0, \varepsilon_{\max}], \ R’ \in  [ R_{\min} , R], \ R \in [R_{\min}, \infty), 
\end{multline*}
where $\varepsilon_{\max}$ and $R_{\min}$ are arbitrary and fixed, and $\omega = R_{\min}/R$.
\end{theorem}
